---
title: "Introduction to IncubationPeriodModels"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to IncubationPeriodModels}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
library(IncubationPeriodModels)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Incubation periods in infectious disease epidemiology

The incubation period is the time between infection and the onset of
symptoms. Understanding this delay is important for many epidemiological
applications including outbreak investigation, transmission modelling,
and estimation of key disease parameters.

Incubation periods are typically positive-valued and right-skewed.
Empirical estimates from outbreak data often show substantial variability
and sometimes heavy tails. As a result, flexible parametric models are
often needed to accurately represent their distribution.

Commonly used models include the log-normal, gamma, and Weibull
distributions. However, these models may not always capture the full
range of shapes observed in incubation period data.

The Burr family of distributions provides a flexible alternative that
can represent a wide range of distributional shapes.

## Burr-family models for incubation periods

The `IncubationPeriodModels` package implements several Burr-family
distributions that can be used to represent incubation period
distributions.

The package currently provides implementations of:

- Burr Type III
- Burr Type X
- Burr Type XII
- A derived Burr-like distribution

These distributions allow flexible modelling of right-skewed delay
distributions and can accommodate varying levels of tail behaviour.

## Simulating incubation period data

To illustrate how these distributions can be used, we simulate
incubation periods from a Burr Type XII distribution.

```{r}
set.seed(123)

incubation <- rburr12(
  n = 1000,
  c = 2,
  k = 3,
  s = 1
)

head(incubation)
```

## Visualising simulated incubation periods

We can visualise the simulated data and compare it with the theoretical
density.

```{r}
hist(
  incubation,
  probability = TRUE,
  breaks = 40,
  main = "Simulated incubation periods",
  xlab = "Time"
)

curve(
  dburr12(x, c = 2, k = 3, s = 1),
  add = TRUE,
  col = "red",
  lwd = 2
)
```

This example illustrates the right-skewed distribution typically
observed for incubation periods.

## Exploring parameter effects

One advantage of Burr-family distributions is that they allow a wide
range of shapes depending on the parameter values.

```{r}
x <- seq(0, 10, length.out = 200)

plot(
  x,
  dburr12(x, c = 1.5, k = 2, s = 1),
  type = "l",
  lwd = 2,
  ylab = "Density",
  xlab = "Time",
  main = "Effect of parameter changes"
)

lines(x, dburr12(x, c = 2, k = 3, s = 1), col = 2, lwd = 2)
lines(x, dburr12(x, c = 3, k = 4, s = 1), col = 3, lwd = 2)
```

Changing the shape parameters modifies the skewness and tail behaviour
of the distribution, allowing different incubation period profiles to
be represented.

## Using distribution functions in modelling

All distributions in the package follow the standard R interface:

| Function | Purpose |
|--------|--------|
| `d*` | density |
| `p*` | cumulative distribution |
| `q*` | quantile |
| `r*` | random generation |

For example:

```{r}
pburr12(5, c = 2, k = 3, s = 1)

qburr12(0.5, c = 2, k = 3, s = 1)
```

These functions make it straightforward to incorporate Burr-family
distributions into simulation studies or likelihood-based inference
frameworks.

## Summary

The Burr family of distributions provides flexible parametric models
for right-skewed delay distributions such as incubation periods.

The `IncubationPeriodModels` package implements several Burr-family
distributions in R, allowing researchers to easily simulate, visualise,
and analyse incubation period data using a consistent interface.
